Inferring generation-interval distributions from contact tracing data: \\ \emph{in prep, PRSB}

### 2018-11-12 doc/network.tex
A few extra comments

\begin{figure}
\includegraphics[width=\textwidth]{../fig/temporal_effect.pdf}
\caption{\textbf{Temporal variation in the mean observed generation interval.}
A deterministic Susceptible-Exposed-Infectious-Recovered (SEIR) model was simulated using Ebola-like parameters and the observed mean generation interval was calculated the course of an epidemic (see methods).
}
\label{fig:censor}
\end{figure}

@@

When generation intervals are sampled through contact tracing, there will be four effects present in the sample: (1) right-censoring effect, (2) egocentric effect, (3) local spatial effect and (4) global spatial effect.
Right-censoring effect and egocentric effect are well-understood.
While local and global spatial effects are difficult to measure, we can make qualitative predictions about their effects on the realized generation intervals and reproductive number. 
Both local and global effects reduce number of infections that occur and shorten generation intervals but the time scale on which they matter differs:
local spatial effect will be important early in an epidemic whereas the global effect will be small until sufficient amount of susceptible individuals become infected.
As a result, both local and egocentric effects affect the initial spread of a disease and therefore the realized reproductive number.

@@

We refer to the first method as the population-level method as it relies on the  observed distribution aggregated across population.
As the observed generation-interval distribution is a weighted distribution of the effective distribution, the effective distribution can be recovered by taking the inverse of the weights, given by \eref{exp}.
Hence, the population-level method relies on the observed generation intervals and the exponential growth rate:
\begin{equation}
g(\tau) \propto \tsub{c}{exp}(\tau) \exp(r\tau).
\end{equation}

While the population-level method is simple and intuitive, it does not use all available information from contact tracing data as it does not take into account who infected whom.
Instead, we can model the initial spread of an epidemic using a poisson process on an individual-level and obtain the following likelihod::
\begin{equation}
\mathcal{L}(\RR, \theta) = \prod_{j=1}^N \left(\RR^{n_j} \exp \left(- \RR \int_0^{\tsub{t}{censor} - t_j} g(s;\theta) ds \right) \prod_{i=1}^{n_j} g(\tau_{i, j}; \theta) \right),
\end{equation}
where $N$ is the total number of infected individuals, $n_j$ is the number of individuals infected by $j$, $\tsub{t}{censor}$ is the time period which censoring was performed until, $t_j$ is the time at which infector $j$ was infected, $\tau_{i,j}$ is the observed generation interval between infector $j$ and infectee $i$, and $\theta$ is the (vector) parameter of the generation-interval distribution $g$.
A similar method was proposed by \cite{forsberg2008likelihood} but it relies on discretized incidence reports rather than the observed generation intervals.

\swp{I might say something about independence transmission events in the results if I decide to compare forsberg2008 as well -- see forsberg 2008...}
\swp{I got stuck on forsberg 2008. I get weird estimates with wide confidence intervals but I feel like that's what their method is supposed to do (based on figure 3 in their paper). It doesn't use generation interval either so I think we can just ignore it?}

>>>>

\begin{figure}
\includegraphics[width=\textwidth]{../fig/temporal_effect.pdf}
\caption{\textbf{Temporal variation in the mean observed generation interval.}
A deterministic Susceptible-Exposed-Infectious-Recovered (SEIR) model was simulated using Ebola-like parameters and the observed mean generation interval was calculated over the course of an epidemic (see methods).
\jd{Meaning, left-censored?}
}
\label{fig:censor}
\end{figure}

@@

When generation intervals are sampled through contact tracing, there will be four effects present in the sample: (1) right-censoring effect, (2) egocentric effect, (3) local spatial effect and (4) global spatial effect.
Right-censoring effect and egocentric effect are well-understood.
While local and global spatial effects are difficult to measure, we can make qualitative predictions about their effects on the realized generation intervals and reproductive number. 
\jd{Any objection to conceptualizing global as mean susceptible depletion? Then it's easy to measure. It's the non-egocentric non-global that's the problem.}
Both local and global effects reduce number of infections that occur and shorten generation intervals but the time scale on which they matter differs:
local spatial effect will be important early in an epidemic whereas the global effect will be small until sufficient amount of susceptible individuals become infected.
As a result, both local and egocentric effects affect the initial spread of a disease and therefore the realized reproductive number.

@@

We refer to the first method as the population-level method as it relies on the  observed distribution aggregated across population.
As the observed generation-interval distribution is a weighted distribution of the effective distribution, the effective distribution can be recovered by taking the inverse of the weights, given by \eref{exp}.
Hence, the population-level method relies on the observed generation intervals and the exponential growth rate:
\begin{equation}
g(\tau) \propto \tsub{c}{exp}(\tau) \exp(r\tau).
\end{equation}
While the population-level method is simple and intuitive, it does not use all available information from contact tracing data as it does not take into account who infected whom.

The individual-level method uses the poisson process to estimate a likelihood for the observations from each individual:
\begin{equation}
\mathcal{L}(\RR, \theta) = \prod_{j=1}^N \left(\RR^{n_j} \exp \left(- \RR \int_0^{\tsub{t}{censor} - t_j} g(s;\theta) ds \right) \prod_{i=1}^{n_j} g(\tau_{i, j}; \theta) \right),
\end{equation}
where $N$ is the total number of infected individuals, $n_j$ is the number of individuals infected by $j$, $\tsub{t}{censor}$ is the time period which censoring was performed until, $t_j$ is the time at which infector $j$ was infected, $\tau_{i,j}$ is the observed generation interval between infector $j$ and infectee $i$, and $\theta$ is the (vector) parameter of the generation-interval distribution $g$.
A similar method was proposed by \cite{forsberg2008likelihood} but it relies on discretized incidence reports rather than the observed generation intervals.

\swp{I might say something about independence transmission events in the results if I decide to compare forsberg2008 as well -- see forsberg 2008...}
\swp{I got stuck on forsberg 2008. I get weird estimates with wide confidence intervals but I feel like that's what their method is supposed to do (based on figure 3 in their paper). It doesn't use generation interval either so I think we can just ignore it?}
\jd{If it really is similar in some important way just stick with what you've already said. Otherwise, ignore it.}
